---
title: "Hotel Booking Demand"
author: "Jing Zhang"
date: "2020/5/7"
output: 
  beamer_presentation:
    colortheme: dolphin
    theme: Madrid
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

## Motivation

* When booking a hotel, people consider many factors, such as the best time to book a hotel, the location of the hotel and so on. For the hotel manager, how to better attract customers to book the hotel?

* Now I want to explore this issue from a different perspective. I want to know how likely it is for people to choose to eat in hotels so as to provide some constructive suggestions for hotels to attract more customers. 

* My final goal is to make predictions about the possibility that people will choose to eat in hotels.

## Data preparation and description

* This dataset comes from the Kaggle and contains booking information for a city hotel and a resort hotel, and includes information such as when the booking was made, length of stay, the number of adults, children, and/or babies, and the number of available parking spaces, among other things. 

```{r}
suppressMessages(library(tibble))
suppressMessages(library(tidyr))
suppressMessages(library(reshape2))
suppressMessages(library(dplyr))
suppressMessages(library(skimr))
suppressMessages(library(nnet))
suppressMessages(library(GGally))
suppressMessages(library(ggplot2))
suppressMessages(library(corrplot))
suppressMessages(library(pROC))
suppressMessages(library(plotROC))
suppressMessages(library(e1071))
suppressMessages(library(rpart))
suppressMessages(library(RColorBrewer))
hotel = read.csv('hotels.csv')
hotel = as_tibble(hotel)
summary(skim(hotel))
```

# Exploratory data analysis

## How the features are correlated to each other

```{r}
## Handle the missing values

#is.na(hotel)
#which(is.na(hotel), arr.ind = T) # column 11 has missing values
#names(hotel)[11]
hotel = replace_na(hotel, list(children = 0))
#which(is.na(hotel), arr.ind = T)


## Handle the columns "meal" and "required_car_parking_spaces"

hotel <- hotel %>%
  mutate(
    meal = case_when(
      meal == "SC" ~ "meal_none",
      meal == "Undefined" ~ "meal_none",
      TRUE ~ as.character(meal)
    ),
    required_car_parking_spaces = case_when(
      required_car_parking_spaces > 0 ~ "parking",
      TRUE ~ "none"
    ),
  )
X_meal = distinct(hotel, meal)

## Set dummy variables based on column "meal"
meal_dummy = as_tibble(class.ind(hotel$meal))
hotel = hotel %>% transform("BB" = meal_dummy$BB) %>%
  transform("FB" = meal_dummy$FB) %>%
  transform("HB" = meal_dummy$HB) %>%
  transform("meal_none" = meal_dummy$meal_none)
  
# dim(hotel)  119390 X 36

## Let’s filter to only the bookings that did not cancel.

hotel_0 = filter(hotel, is_canceled == 0)

## Let's plot how the features are correlated to each other.

hotel_numeric = tibble(zero = rep(0, nrow(hotel_0)))
temp_name = c("zero")
for(i in 1:ncol(hotel_0))
{
  if(class(hotel_0[[i]]) == "integer" | class(hotel_0[[i]]) == "numeric")
  {
    temp = unlist(select(hotel_0, i))
    temp_name = c(temp_name, names(select(hotel_0, i)))
    hotel_numeric = add_column(hotel_numeric, temp)
    names(hotel_numeric) = paste("X", 1:ncol(hotel_numeric),sep = "")
  }
}
names(hotel_numeric) = temp_name
# dim(hotel_numeric) ## 75166    18

hotel_numeric = select(hotel_numeric, -c(zero, is_canceled))
```

```{r}
## plot heatmap
cor_hotel = as.data.frame(cor(hotel_numeric))
cor_hotel$ID = rownames(cor_hotel)
cor_hotel_0 = melt(cor_hotel, id.vars = c("ID"))
#view(cor_hotel_0)
p <- ggplot(cor_hotel_0, aes(x = variable, y = ID, fill = value))+
  theme(panel.background = element_blank(),axis.line = element_line(colour="black"))+
  theme(axis.ticks.x = element_blank()) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(legend.title = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(plot.margin = unit(rep(3, 4), 'lines'))
p + geom_raster()
```

## Do elders with children choose to eat in hotels?

```{r}
ggplot(hotel_0, aes(x = children)) + 
  geom_bar(aes(fill = factor(meal)), position="fill") +
  labs(y = "proportion") +
  theme(plot.margin = unit(rep(3, 4), 'lines'))
```


## Do elders with babies choose to eat in hotels?

```{r}
ggplot(hotel_0, aes(x = babies)) + 
  geom_bar(aes(fill = factor(meal)), position="fill") +
  labs(y = "proportion") +
  scale_x_continuous(breaks = seq(min(hotel_0$babies), max(hotel_0$babies), 1)) +
  theme(plot.margin = unit(rep(3, 4), 'lines'))
```


## How about availability of parking?
```{r}
hotel_0 %>%
  count(hotel, required_car_parking_spaces, meal) %>%
  group_by(hotel, meal) %>%
  mutate(proportion = n / sum(n)) %>%
  ggplot(aes(required_car_parking_spaces, proportion, fill = meal)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format()) +
  facet_wrap(~hotel, nrow = 2) +
  labs(x = NULL, y = "Proportion of hotel stays") +
  theme(plot.margin = unit(rep(0.5, 4), 'lines'))
```

## Does the average daily rate affect people's choice to eat in hotels?

```{r}
ggplot(hotel_0, aes(x = meal, y = adr)) + 
  geom_boxplot(fill = brewer.pal(7, "Blues")[4], color = brewer.pal(7, "Blues")[6]) +
  labs(y = "Average Daily Rate") +
  theme(plot.margin = unit(rep(3, 4), 'lines'))
```


# Formal Modeling

```{r}
hotel = hotel %>% mutate(
  children = case_when(
    children > 0 ~ "children",
    TRUE ~ "none"
  )
)
#distinct(hotel, children)

# "meal_none" <- 0
# "HB" <- 1
# "FB" <- 1
# "BB" <- 1

temp_meal = sub("meal_none", "0", hotel$meal) 
temp_meal = sub("HB", "1", temp_meal) 
temp_meal = sub("FB", "1", temp_meal) 
temp_meal = sub("BB", "1", temp_meal)
temp_meal = as.numeric(temp_meal)
hotel = add_column(hotel, temp_meal)
hotel = rename(hotel, order_meal = temp_meal)
# dim(hotel) 119390 X 37

## Dealing with maxima and minima

a = hotel # a is not a tibble
q20 = quantile(a$adr, 0.2)
q80 = quantile(a$adr, 0.8)
a[a$adr < q20, ]$adr <- q20
a[a$adr > q80, ]$adr <- q80
#summary(a$adr)

hotel_new = as_tibble(a)
names_new = names(hotel)
names(hotel_new) <- names_new
# dim(hotel_new) 119390 X 37
#head(hotel_new)

## Choose the features
hotel_new = filter(hotel_new, is_canceled == 0)
hotel_f = select(hotel_new, children, babies, required_car_parking_spaces, adr, order_meal)
# dim(hotel_f) 75166 X 5


## Divide the dataset into training set and test set.

set.seed(123)
train_data = sample_frac(hotel_f, 0.8)
test_data = sample_frac(hotel_f, 0.2)
```

## Logistic Regression

```{r}
glm_log = glm(order_meal~., data = train_data, family = binomial("logit"))
#glm_log
#summary(glm_log)

#step_log = step(glm_log, direction = "both")
#summary(step_log)

### Test the model
pre_log = predict(glm_log, test_data, type = "response")
#head(pre_log)
fitted_log = ifelse(pre_log > 0.5, 1, 0)
accuracy_log = table(fitted_log, test_data$order_meal)
ggplot(test_data, aes(d = order_meal, m = pre_log)) + geom_roc()
```

## Model

$$\hat{h}_\theta(z) = \frac{1}{1 + \exp{(-z)}}$$
$$z = \hat{\theta}^{T}x$$
$$\hat{\theta} = (3.684, -1.366, 0.433, 1.062, -0.003)^{T}$$
$$x = (x_0, x_{children}, x_{babies}, x_{park}, x_{adr})^{T}$$

## Future Analysis

* Use different methods to make predictions, such as decision tree, SVM and so on.

* Compare the used methods and analyze their pros and cons.

* Choose the best model to make predictions.

## The end

\begin{center}
Thank you!
\end{center}

